---
title: "Lrn Incercept MLE"
output: html_document
---

Point to correct package library - required on my computer due to R upgrade

```{r}
.libPaths("/home/david/R/x86_64-pc-linux-gnu-library/4.2")
```

The purpose of this script is to:
Import MOFA model inferred from learning set
Extract or calculate learning set intercepts
For Guassian data the intercepts are available from the MOFA model
For other data they are calculated with MLE

```{r}
library(ggplot2)
library(MOFA2)
library(rhdf5)
library(rjson)
```

Specify which simulation and factorization configurations to import
The script loops through the configurations and within each configuration it loops through the simulation instances.

```{r}
BaseDir = getwd() # or change it to somewhere else

Configs = c(
  "SimConfig1"
  )

scale_views = 'False'
center_groups = 'True'
FactorSelection = 'ActiveK'
```

Loop through configurations, simulations and learning set factorizations
As simulation and factorization was done in python the indices start at 0

```{r}
for (Config in Configs){
  
  print(paste0(Config,":",Sys.time()))

  # Specify global paths
  
  InputDir = file.path(BaseDir, Config)

  FctrznsDir = file.path(InputDir,
                         paste0("Fctrzn_Lrn","_scale_",scale_views,"_center_",center_groups,"_",FactorSelection))
  
  # Import metadata related to simulations and factorizations
  
  SimMeta = fromJSON(file = file.path(InputDir,"sim_meta.json"))
  M = length(SimMeta$D)
  Sims = SimMeta$Sims
  
  for (sim in 1:Sims-1){

    fit_start_time = Sys.time()
  
    # simulation, factorization and transfer learning directories
    SimDir = file.path(InputDir,"SimData",paste0("sim",sim))
    FctrznDir = file.path(FctrznsDir,paste0("Fctrzn",sim))
    
    # Load the model and information about what features were used in factorization
    
    InputModel = file.path(FctrznDir,"Model.hdf5")
    Fctrzn = load_model(file = InputModel)
    likelihoods = as.vector(Fctrzn@model_options$likelihoods)
    
    ####
    ## LOOP THROUGH LIKELIHOODS TO EXTRACT OR CALCULATE INTERCEPTS AND SDs
    ####
    
    InterceptsNaive = vector("list")
    Intercepts = vector("list")
    InterceptsMethod = vector("list")
    
    for (m in 1:M){

      InterceptsMLE = numeric()
      InterceptsMethodTmp = character()
      YLrn = t(Fctrzn@data[[m]]$group0)
      ZWLrn = Fctrzn@expectations$Z$group0 %*% t(Fctrzn@expectations$W[[m]])
      DLrn = dim(YLrn)[2]
    
      if (likelihoods[m]=="gaussian"){
        
        ## gaussian intercepts are already available
        InterceptsNaive[[m]] = as.vector(Fctrzn@intercepts[[m]]$group0)
        Intercepts[[m]] = InterceptsNaive[[m]]
        InterceptsMethod[[m]] = rep("Naive",length(Intercepts[[m]]))
        
      } else if (likelihoods[m]=="poisson"){
        
        ## naive intercept based on approximation to feature means of ZW
        InterceptsNaive[[m]] = as.vector(log(-1 + exp(colMeans(YLrn))))
        
        ## mle estimate of intercept for ZW
        ## if optimiser fails for a feature will return the naive estimate
        
        for (d in 1:DLrn){

          # nLL = function(interceptMLE) -sum(stats::dpois(YLrn[,d], log(1 + exp(ZWLrn[,d]+interceptMLE)), log = TRUE))
          
          nLL = function(interceptMLE) -sum(log(
            stats::dpois(YLrn[,d], log(1 + exp(ZWLrn[,d]+interceptMLE)))[stats::dpois(YLrn[,d], log(1 + exp(ZWLrn[,d]+interceptMLE)))!=0]
            ))
          
          # interceptMLEfit = try(as.vector(stats4::mle(nLL, start=list(interceptMLE=0))@coef[1]))
          
          interceptMLEfit = try(as.vector(
            stats4::mle(nLL, start=list(interceptMLE=InterceptsNaive[[m]][d]))@coef[1]
            ))
          
          if (class(interceptMLEfit)=="try-error"){
            InterceptsMLE = c(InterceptsMLE,InterceptsNaive[[m]][d])
            InterceptsMethodTmp = c(InterceptsMethodTmp,"Naive")
          } else {
            InterceptsMLE = c(InterceptsMLE,interceptMLEfit)
            InterceptsMethodTmp = c(InterceptsMethodTmp,"MLE")
            }
        }
        
        Intercepts[[m]] = InterceptsMLE
        InterceptsMethod[[m]] = InterceptsMethodTmp
        
      } else if (likelihoods[m]=="bernoulli"){
        
        ## naive intercept based on approximation to feature means of ZW
        InterceptsNaive[[m]] = log(
          as.vector(colMeans(YLrn))/(1-as.vector(colMeans(YLrn)))
          )
        
        ## mle estimate of intercept for ZW
        ## if optimiser fails for a feature will return the naive estimate
        
        for (d in 1:DLrn){

          # nLL = function(interceptMLE) -sum(stats::dbinom(YLrn[,d], size=1, plogis(ZWLrn[,d]+interceptMLE), log = TRUE))
          
          nLL = function(interceptMLE) -sum(log(
            stats::dbinom(YLrn[,d], size=1, plogis(ZWLrn[,d]+interceptMLE))[stats::dbinom(YLrn[,d], size=1, plogis(ZWLrn[,d]+interceptMLE))!=0]
            ))
          
          
          # interceptMLEfit = try(as.vector(stats4::mle(nLL, start=list(interceptMLE=0))@coef[1]))
          
          interceptMLEfit = try(as.vector(
            stats4::mle(nLL, start=list(interceptMLE=InterceptsNaive[[m]][d]))@coef[1]
            ))
          
          if (class(interceptMLEfit)=="try-error"){
            InterceptsMLE = c(InterceptsMLE,InterceptsNaive[[m]][d])
            InterceptsMethodTmp = c(InterceptsMethodTmp,"Naive")
          } else {
            InterceptsMLE = c(InterceptsMLE,interceptMLEfit)
            InterceptsMethodTmp = c(InterceptsMethodTmp,"MLE")
            }
        }
        
        Intercepts[[m]] = InterceptsMLE
        InterceptsMethod[[m]] = InterceptsMethodTmp
        
      } else {
        
        InterceptsNaive[[m]] = numeric()
        Intercepts[[m]] = numeric()
        InterceptsMethod[[m]] = character()
        
      }
    }
    
    ## save the intercepts in the relevant factorization folder
    
    fit_end_time = Sys.time()
    
    EstimatedIntercepts = list(
      InterceptsNaive = InterceptsNaive,
      Intercepts = Intercepts,
      InterceptsMethod = InterceptsMethod,
      likelihoods = likelihoods,
      fit_start_time = fit_start_time,
      fit_end_time = fit_end_time
    )
    
    saveRDS(EstimatedIntercepts,file.path(FctrznDir,"EstimatedIntercepts.rds"))
    
  }
}
```


