---
title: "FctrznEvaluation"
output: html_document
---

This script is for evaluating factorizations and comparing them with respect to uncovering underlying matrices. It loops through simulation configurations and instances.
For each configuration it outputs an rds file with evaluation metrics

How to evaluate?

 - test Z scores between target set clusters
 - identify scores with statistically significant differences
 - find groundtruth weight vectors that resemble the inferred weight vectors associated with the differentially active scores
 - are the groundtruth scores associated with these weight vectors actually differentially active?
 
Required Packages

```{r}
library(ggplot2)
library(MOFA2)
library(rhdf5)
library(rjson)
library(gridExtra)
library(scales)
library(knitr)
library(kableExtra)
library(mogsa)
```


 
Global specifics for importing
 
```{r}
Seed = 1234567
mode(Seed) = 'integer'
set.seed(Seed)

BaseDir = getwd() # or change it to somewhere else

Configs = c(
  "SimConfig3",
  "SimConfig3K",
  "SimConfig3Zsd",
  "SimConfig4",
  "SimConfig4K",
  "SimConfig4Zsd"
  )

scale_views = 'False'
center_groups = 'True'
FactorSelection = 'ActiveK'
FctrznMethods = c('Direct_Trg', 'TL_VI', 'Random','IntNMF', 'MoCluster',
                  'TL_VI_permute_0', 'TL_VI_permute_10', 'TL_VI_permute_20', 'TL_VI_permute_30',
                  'TL_VI_permute_40', 'TL_VI_permute_50', 'TL_VI_permute_60', 'TL_VI_permute_70',
                  'TL_VI_permute_80', 'TL_VI_permute_90', 'TL_VI_permute_100'
                  )

FeatWseScl = TRUE ## if TRUE then weight vectors scaled featurewise, otherwise factorwise
# ScaleW = function(x) {x/max(abs(x),na.rm = TRUE)} ## function for scaling weight vectors
ScaleW = function(x) {x/norm(matrix(x),type="F")}
pvalueTH = 0.05 ## THRESHOLD FOR DIFFERENTIALLY ACTIVE
MltTst = FALSE ## whether to use multiple testing adjustments to p-values - not needed here as using the groundtruth factor score means to check against

CorMethod = "pearson" ## either "pearson" or "spearman" - for correlations between w vectors
corCutOff = 0.05 ## pvalue threshold for whether a correlation is significant

```

loop through configurations and simulated data sets

```{r}
for (Config in Configs){
  
  print(Config)
  # Directories for simulated data and different factorization types
  
  InputDir = file.path(BaseDir, Config)
  SimsDir = file.path(InputDir,"SimData")

  # Import metadata related to simulations and factorizations
  
  SimMeta = fromJSON(file = file.path(InputDir,"sim_meta.json"))
  M = length(SimMeta$D)
  YDists = SimMeta$YDists
  Sims = SimMeta$Sims
  CLrn = SimMeta$CLrn
  CTrgN = SimMeta$CTrgN
  CTrg1 = (1:(CTrgN*2))<=CTrgN
  CTrg2 = !CTrg1
  
  # set up vectors to store results
  
  sim_iteration = integer()
  fctrzn_type = character()
  score_name = character()
  score = numeric()
  
  for (sim in 1:Sims-1){
    
    print(paste0("sim",sim))
    
    # directory for each simulation
    SimDir = file.path(SimsDir,paste0("sim",sim))
    
    # import groundtruth weights
    WGroundtruth = vector("list")
    for (m in 1:M-1){
      WGroundtruth[[m+1]] = as.matrix(read.csv(file.path(SimDir,paste0("W",m,".csv")),header = FALSE))
    }
    
    # import groundtruth clusterwise score means
    ZGroundtruth = as.matrix(read.csv(file.path(SimDir,paste0("ZMuC.csv")),header = FALSE))
    ZGroundtruth = ZGroundtruth[1:dim(ZGroundtruth)[1] > CLrn,]
    
    ## groundtruth score matrix perturbed scores - between trg set clusters
      
    GT_SigFct = as.vector(which(ZGroundtruth[1,]!=ZGroundtruth[2,]))
    GT_positives = length(GT_SigFct)
    
    sim_iteration = c(sim_iteration,sim)
    fctrzn_type = c(fctrzn_type,"GroundTruth")
    score_name = c(score_name,"GT_positives")
    score = c(score,GT_positives)
    
    # info about which set the samples belong to
    IfLrn = read.csv(file.path(SimDir,"IfLrn.csv"),header=FALSE)
    IfLrn = as.vector(IfLrn$V1)
    
    # import data for the Trg Fctrzn and evaluate against ground truth
    for (fm in FctrznMethods){
      
      # set directory for factorization of target and import
      
      if (fm=='Direct_Trg'){
        
        FctrznsTrgDir = file.path(InputDir,
                         paste0("Fctrzn_Trg_scale_",scale_views,"_center_",center_groups,"_",FactorSelection))
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("Fctrzn",sim))
        
        FctrznTrg = load_model(file = file.path(FctrznTrgDir,"Model.hdf5"))
        FTrgVarKp = vector("list")
        for (m in 1:M-1){
          FTrgVarKp[[m+1]] = as.vector(
            read.csv(file.path(FctrznTrgDir,paste0("FVarKp_m",m,"g_0.csv")),header = FALSE)$V1
            )
        }
        FctrznTrg_W_Fit = FctrznTrg@expectations$W
        FctrznTrg_Z = FctrznTrg@expectations$Z$group0
        FitSeconds = sum(FctrznTrg@training_stats$time)
        
      } else if (fm=='Direct_All'){
        
        FctrznsTrgDir = file.path(InputDir,
                           paste0("Fctrzn_All_scale_",
                                  scale_views,"_center_",center_groups,"_",
                                  FactorSelection))
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("Fctrzn",sim))
        
        FctrznTrg = load_model(file = file.path(FctrznTrgDir,"Model.hdf5"))
        FTrgVarKp = vector("list")
        for (m in 1:M-1){
          FTrgVarKp[[m+1]] = as.vector(
            read.csv(file.path(FctrznTrgDir,paste0("FVarKp_m",m,"g_0.csv")),header = FALSE)$V1
            )
        }
        FctrznTrg_W_Fit = FctrznTrg@expectations$W
        FctrznTrg_Z = FctrznTrg@expectations$Z$group0[IfLrn==0,]
        FitSeconds = sum(FctrznTrg@training_stats$time)
      
        } else if (substr(fm,1,5)=='TL_VI'){  
        
        FctrznsTrgDir = file.path(InputDir,
                         paste0("TL","_scale_",
                                scale_views,"_center_",center_groups,"_",
                                FactorSelection,substr(fm,3,nchar(fm))))
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("TL",sim))
        
        FctrznTrg = readRDS(file.path(FctrznTrgDir,"TL_VI_data.rds"))
        FTrgVarKp = FctrznTrg$TrgKp
        FctrznTrg_W_Fit = FctrznTrg$Fctrzn_Lrn_W_fltrd
        FctrznTrg_Z = FctrznTrg$ZMu
        FitSeconds = as.numeric(difftime(FctrznTrg$fit_end_time,FctrznTrg$fit_start_time, units="secs"))
        
        
        } else if (fm=='IntNMF'){  
        
        FctrznsTrgDir = file.path(InputDir,"Fctrzn_Trg_IntNMF")
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("Fctrzn",sim))
        
        FctrznTrg = readRDS(file.path(FctrznTrgDir,"IntNMF_data.rds"))
        FTrgVarKp = FctrznTrg$TrgKp
        FctrznTrg_W_Fit = FctrznTrg$IntNMF_fctrzn$H
        for (m in 1:M){
          FctrznTrg_W_Fit[[m]] = t(FctrznTrg_W_Fit[[m]])
        }
        FctrznTrg_Z = FctrznTrg$IntNMF_fctrzn$W
        FitSeconds = as.numeric(difftime(FctrznTrg$fit_end_time,FctrznTrg$fit_start_time, units="secs"))
        
        
        } else if (fm=='MoCluster'){  
        
        FctrznsTrgDir = file.path(InputDir,"Fctrzn_Trg_MoCluster")
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("Fctrzn",sim))
        
        FctrznTrg = readRDS(file.path(FctrznTrgDir,"MoCluster_data.rds"))
        FTrgVarKp = FctrznTrg$TrgKp
        DinM = as.vector(FctrznTrg$MoCluster_fctrzn@tab.dim[1,])
        DinMCumul = cumsum(DinM)
        FctrznTrg_W_Fit = vector("list")
        for (m in 1:M){
          FctrznTrg_W_Fit[[m]] = moaCoef(FctrznTrg$MoCluster_fctrzn)$coefMat[(DinMCumul[m]-DinM[m]+1):(DinMCumul[m]),]
        }
        FctrznTrg_Z = moaScore(FctrznTrg$MoCluster_fctrzn)
        FitSeconds = as.numeric(difftime(FctrznTrg$fit_end_time,FctrznTrg$fit_start_time, units="secs"))
        
        
      } else if (fm=='Random'){
        
        FctrznsTrgDir = file.path(InputDir,
                         paste0("TL","_scale_",
                                scale_views,"_center_",center_groups,"_",
                                FactorSelection,"_VI"))
        FctrznTrgDir = file.path(FctrznsTrgDir,paste0("TL",sim))
        
        FctrznTrg = readRDS(file.path(FctrznTrgDir,"TL_VI_data.rds"))
        FTrgVarKp = FctrznTrg$TrgKp
        FctrznTrg_W_Fit = FctrznTrg$Fctrzn_Lrn_W_fltrd
        FctrznTrg$fit_start_time = Sys.time()
        FctrznTrg_Z = FctrznTrg$ZMu
        
        # randomize Z
        FctrznTrg_Z = matrix(
          rnorm(length(as.vector(FctrznTrg_Z)),0,1),
          nrow=dim(FctrznTrg_Z)[1], ncol=dim(FctrznTrg_Z)[2]
        )
        
        FctrznTrg$fit_end_time = Sys.time()
        FitSeconds = as.numeric(difftime(FctrznTrg$fit_end_time,FctrznTrg$fit_start_time, units="secs"))
        
      }
        
      ## scaled groundtruth weights relevant to target set factorization
      FctrznTrg_W_GT = vector("list")
      for (m in 1:M){
        
        if (fm=='Random'){
          FTrgVarKp[[m]] = sample(FTrgVarKp[[m]],length(FTrgVarKp[[m]]),replace = FALSE)
        }
        
        if (FeatWseScl){
          FctrznTrg_W_GT[[m]] = t(apply(WGroundtruth[[m]][FTrgVarKp[[m]]==1,], 1, ScaleW))
        } else {
          FctrznTrg_W_GT[[m]] = apply(WGroundtruth[[m]][FTrgVarKp[[m]]==1,], 2, ScaleW)
        }
        FctrznTrg_W_GT[[m]][is.na(FctrznTrg_W_GT[[m]])]=0
        FctrznTrg_W_GT[[m]] = base::scale(FctrznTrg_W_GT[[m]], center = TRUE, scale = FALSE)
      }
      FctrznTrg_W_GT_Ccnt = do.call(rbind,FctrznTrg_W_GT)
      FctrznTrg_W_GT_Ccnt[is.nan(FctrznTrg_W_GT_Ccnt)] = 0
      
      ## scaled weights inferred by target set factorization
      for (m in 1:M){
        
        if (fm=='Random'){
          FctrznTrg_W_Fit[[m]] = matrix(
          rnorm(length(as.vector(FctrznTrg_W_Fit[[m]])),0,1),
          nrow=dim(FctrznTrg_W_Fit[[m]])[1], ncol=dim(FctrznTrg_W_Fit[[m]])[2]
          )
        }
        
        if (FeatWseScl){
          FctrznTrg_W_Fit[[m]] = t(apply(FctrznTrg_W_Fit[[m]],1, ScaleW))
        } else {
          FctrznTrg_W_Fit[[m]] = apply(FctrznTrg_W_Fit[[m]],2, ScaleW)
        }
        FctrznTrg_W_Fit[[m]][is.na(FctrznTrg_W_Fit[[m]])]=0
        FctrznTrg_W_Fit[[m]] = base::scale(FctrznTrg_W_Fit[[m]], center = TRUE, scale = FALSE)
      }
      FctrznTrg_W_Fit_Ccnt = do.call(rbind,FctrznTrg_W_Fit)
      FctrznTrg_W_Fit_Ccnt[is.nan(FctrznTrg_W_Fit_Ccnt)] = 0
      
      #### Check for significant scores from Z matrix
      
      FctrznTrg_Z_pv = numeric()
      for (f in 1:dim(FctrznTrg_Z)[2]){
        WRSTest = wilcox.test(x=FctrznTrg_Z[CTrg1,f], y = FctrznTrg_Z[CTrg2,f])
        FctrznTrg_Z_pv = c(FctrznTrg_Z_pv,WRSTest$p.value)
      }
      if (MltTst){
        FctrznTrg_Z_pv = p.adjust(FctrznTrg_Z_pv, method="BH")
      }
      
      #### Find associated weight vectors and predicted factors
      # for each differently expressed z factor - what is the most correlated groundtruth w factor?
      
      FctrznTrg_SigFct = integer()
      if (sum(FctrznTrg_Z_pv<pvalueTH)==0){
        FctrznTrg_SigFct = c(FctrznTrg_SigFct,0)
      } else {
          for (f in which(FctrznTrg_Z_pv<pvalueTH)){
            FctrCors = numeric()
            FctrCorsPVs = numeric()
            for (fGT in 1:dim(FctrznTrg_W_GT_Ccnt)[2]){
              FctrCors = c(FctrCors,
                           cor(x=FctrznTrg_W_Fit_Ccnt[,f], 
                               y = FctrznTrg_W_GT_Ccnt[,fGT],
                               method = CorMethod)
                           )
              FctrCorsPVs = c(FctrCorsPVs,
                              cor.test(x=FctrznTrg_W_Fit_Ccnt[,f],
                                       y = FctrznTrg_W_GT_Ccnt[,fGT],
                                       alternative = c("two.sided"),
                                       method = CorMethod)$p.value
                              )
              
            }
            FctrCorsNums = 1:length(FctrCors)
            # only statistically significant correlations considered
            FctrCors = FctrCors[FctrCorsPVs<corCutOff]
            FctrCorsNums = FctrCorsNums[FctrCorsPVs<corCutOff]
            if(length(FctrCors)>=1){
              FctrCorsRks = base::rank(-FctrCors, ties="first")
              FctrznTrg_SigFct = c(FctrznTrg_SigFct, FctrCorsNums[FctrCorsRks == 1])
            } else {
              FctrznTrg_SigFct = c(FctrznTrg_SigFct,0)
            }
          }
      }
      FctrznTrg_SigFct = unique(FctrznTrg_SigFct[FctrznTrg_SigFct>0])
      
      ## scores for evaluating the  factorization method
      
      Inf_positives = length(FctrznTrg_SigFct)
      True_positives = sum(is.element(FctrznTrg_SigFct,GT_SigFct))
      if (Inf_positives==0){
        Precision=0
      } else {
        Precision = True_positives/Inf_positives
      }
      if (GT_positives==0){
        Recall=0
      } else {
        Recall = True_positives/GT_positives
      }
      if ((Precision+Recall)==0){
        F1=0
      } else {
        F1 = (2*Precision*Recall)/(Precision+Recall)
      }
      
      sim_iteration = c(sim_iteration,rep(sim,6))
      fctrzn_type = c(fctrzn_type,rep(fm,6))
      score_name = c(score_name,c("Inf_positives","True_positives","Precision","Recall","F1","FitSeconds"))
      score = c(score,c(Inf_positives,True_positives,Precision,Recall,F1,FitSeconds))
      
      ## remove before replacing
      rm(list=c('FctrznsTrgDir', 'FctrznTrgDir',
                'FctrznTrg', 'FTrgVarKp', 'FctrznTrg_W_Fit', 'FctrznTrg_Z', 'FitSeconds'))
      invisible(gc())
    
    }
    
    ## remove before replacing
    rm(list=c('SimDir','WGroundtruth','ZGroundtruth','GT_SigFct','GT_positives','IfLrn'))
    invisible(gc())
      
  }
  
  FctrznEvaluation = data.frame(
    sim_iteration = sim_iteration,
    fctrzn_type = fctrzn_type,
    score_name = score_name,
    score = score)
  
  FctrznEvaluationLs = list(
  "FctrznEvaluation" = FctrznEvaluation,
  "FctrznMethods" = FctrznMethods,
  "Seed" = Seed,
  "pvalueTH" = pvalueTH,
  "FeatWseScl" = FeatWseScl,
  "MltTst" = MltTst,
  "CorMethod" = CorMethod,
  "corCutOff" = corCutOff
  )

  # export the data
  saveRDS(FctrznEvaluationLs,file.path(InputDir,"FctrznEvaluations.rds"))
  
  ## remove before replacing
  rm(list=c('InputDir','SimsDir','SimMeta','M','YDists','Sims','CLrn','CTrgN','CTrg1','CTrg2',
            'FctrznEvaluation', 'FctrznEvaluationLs'))
  invisible(gc())

}

print("finished")
```






 