---
title: "FctrznEvaluationCmbnd"
output: html_document
---


Import results from various configurations to produce summary plots and tables
Also there is code at the bottom to check for statistically significant differences in evaluation scores

Required Packages

```{r}
library(ggplot2)
library(MOFA2)
library(rhdf5)
library(rjson)
library(gridExtra)
library(scales)
library(knitr)
library(kableExtra)
library(nlme)
library(emmeans)
library(RColorBrewer)

```

Results to import

```{r}

BaseDir = getwd() # or change it to somewhere else

Configs = c(
  "SimConfig3",
  "SimConfig3K",
  "SimConfig3Zsd",
  "SimConfig4",
  "SimConfig4K",
  "SimConfig4Zsd"
  )

```


Loop and import

```{r}
sim_metas = vector("list")
Evaluations = vector("list")

for (i in 1:length(Configs)){
 
  InputDir = file.path(BaseDir, Configs[i])
  
  sim_metas[[i]] = fromJSON(file = file.path(InputDir,"sim_meta.json"))
  
  Evaluations[[i]] = readRDS(file.path(InputDir,"FctrznEvaluations.rds"))
  
}

```


Loop and combine metadata with evaluation scores
and adjust evaluation scores to change NAs to zero


```{r}

for (i in 1:length(Configs)){

  n = dim(Evaluations[[i]]$FctrznEvaluation)[1]
  
  Evaluations[[i]]$FctrznEvaluation$score_adj = Evaluations[[i]]$FctrznEvaluation$score
  Evaluations[[i]]$FctrznEvaluation$score_adj[is.na(Evaluations[[i]]$FctrznEvaluation$score_adj)] = 0
  
  Evaluations[[i]]$FctrznEvaluation$SimConfig = rep(Configs[i],n)
  
  Evaluations[[i]]$FctrznEvaluation$CLrn = rep(sim_metas[[i]]$CLrn,n)
  Evaluations[[i]]$FctrznEvaluation$CLrnNPool = rep(paste0("{",paste(sim_metas[[i]]$CLrnNPool, collapse=","),"}"),n)
  Evaluations[[i]]$FctrznEvaluation$K = rep(sim_metas[[i]]$K,n)
  Evaluations[[i]]$FctrznEvaluation$ZShiftP = rep(sim_metas[[i]]$ZShiftP,n)
  Evaluations[[i]]$FctrznEvaluation$Zsd = rep(sim_metas[[i]]$ZSdRange[1],n)
  
  # Evaluations[[i]]$FctrznEvaluation$D = rep(paste0("{",paste(sim_metas[[i]]$D, collapse=","),"}"),n)
  
  Evaluations[[i]]$FctrznEvaluation$D = rep(sim_metas[[i]]$D[1],n)
  Evaluations[[i]]$FctrznEvaluation$YDists = rep(paste0("{",paste(sim_metas[[i]]$YDists, collapse=","),"}"),n)
  Evaluations[[i]]$FctrznEvaluation$M = rep(length(sim_metas[[i]]$D),n)
  
  Evaluations[[i]]$FctrznEvaluation$pvalueTH = rep(Evaluations[[i]]$pvalueTH,n)
  
  Evaluations[[i]] = Evaluations[[i]]$FctrznEvaluation
  
}

```


Create a single data frame to store the results

```{r}
EvaluationsDF = do.call(rbind,Evaluations)

## check name sof methods - change any for plotting?
table(EvaluationsDF$fctrzn_type)

```

Change names if necessary

```{r}
EvaluationsDF$fctrzn_type[EvaluationsDF$fctrzn_type=="TL_VI"]="MOTL"
EvaluationsDF$fctrzn_type[EvaluationsDF$fctrzn_type=="Direct_Trg"]="Direct"
EvaluationsDF$fctrzn_type[EvaluationsDF$fctrzn_type=="Random"]="Random_TL"

table(EvaluationsDF$fctrzn_type)
```


Boxplots for F1 scores - all configs

```{r}
PlotScore = "F1"

DFToPlot = EvaluationsDF[EvaluationsDF$score_name==PlotScore 
                         # & is.element(EvaluationsDF$fctrzn_type, c("Direct", "MOTL", "IntNMF", "MoCluster" ))
                         & is.element(EvaluationsDF$fctrzn_type, c("Direct", "MOTL" ))
                         & EvaluationsDF$M == 3
                         & EvaluationsDF$D==2000
                         ,]

DFToPlot$fctrzn_type = factor(DFToPlot$fctrzn_type, levels = c("Direct", "MOTL", "IntNMF", "MoCluster"))

DFToPlot$D = paste0('D: ',DFToPlot$D)
DFToPlot$K = paste0('K: ',DFToPlot$K)
DFToPlot$Zsd = paste0('sd: ',format(DFToPlot$Zsd,nsmall=1))
DFToPlot$CLrn = paste0('Learning Groups: ',DFToPlot$CLrn)

ggplot(DFToPlot, aes(x=fctrzn_type, y=score_adj, fill=fctrzn_type))+ 
  ylab(PlotScore) +
  xlab("") +
  # geom_violin() +
  geom_boxplot(outlier.size = 1, outlier.shape = 1, fatten=1) +
  stat_summary(fun=mean, geom="point", shape=18, color="black", size=2) +
  facet_grid(K+Zsd~CLrn, scales = 'free_y') +
  coord_flip()+
  theme_bw()+
  theme(
    axis.title.x = element_text(size=6),
    axis.text.x = element_text(size=6),
    axis.text.y = element_text(size=6),
    strip.text.x = element_text(size=6),
    strip.text.y = element_text(angle = 0, size=6),
    legend.position="none",
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank()
    )+
  scale_fill_brewer(palette = "Set2")
ggsave(
  file.path(BaseDir,paste0(PlotScore,"Plots.pdf")),
       width = 15, height = 10, units = "cm"
  )
ggsave(
  file.path(BaseDir,paste0(PlotScore,"Plots.png")),
       width = 15, height = 10, units = "cm"
  )
```

Line graphs of average F1 scores for MOTL based on permuted W matrices

```{r}
PlotScore = "F1"

DFToPlot = EvaluationsDF[EvaluationsDF$score_name==PlotScore 
                         & is.element(EvaluationsDF$fctrzn_type, 
                                      c("Direct",
                                        "TL_VI_permute_0", "TL_VI_permute_10", "TL_VI_permute_20",
                                        "TL_VI_permute_30", "TL_VI_permute_40", "TL_VI_permute_50",
                                        "TL_VI_permute_60", "TL_VI_permute_70", "TL_VI_permute_80",
                                        "TL_VI_permute_90", "TL_VI_permute_100"
                                        ))
                         & EvaluationsDF$M == 3
                         & EvaluationsDF$D==2000
                         ,]

DFToPlot$D = paste0('D: ',DFToPlot$D)
DFToPlot$K = paste0('K: ',DFToPlot$K)
DFToPlot$Zsd = paste0('sd: ',format(DFToPlot$Zsd,nsmall=1))
DFToPlot$CLrn = paste0('Learning Groups: ',DFToPlot$CLrn)

DFToPlot = do.call(data.frame, 
                       aggregate(score_adj ~ K+Zsd+CLrn+fctrzn_type, 
                               DFToPlot,
                               FUN = "mean")
                     )

DFToPlot_TL = DFToPlot[substr(DFToPlot$fctrzn_type,1,13)=="TL_VI_permute",]
DFToPlot_TL$Permuted = as.numeric(substr(DFToPlot_TL$fctrzn_type,15,nchar(DFToPlot_TL$fctrzn_type)))/100
DFToPlot_TL$fctrzn_type = 'MOTL'

DFToPlot_Direct = DFToPlot[DFToPlot$fctrzn_type=="Direct",]

DFToPlot <- merge(x=DFToPlot_TL, 
                  y=DFToPlot_Direct, 
                  by=c("K","Zsd", "CLrn"), all.x=TRUE)

ggplot(DFToPlot, aes(x=Permuted, y=score_adj.x, colour = fctrzn_type.x))+ 
  ylab("Average F1 score") +
  xlab("Proportion of feature vectors permuted") +
  geom_point(color = RColorBrewer::brewer.pal(3, "Set2")[2]) +
  geom_line(color = RColorBrewer::brewer.pal(3, "Set2")[2], linewidth=0.8) +
  facet_grid(K+Zsd~CLrn, scales = 'fixed') +
  theme_bw()+
  theme(
    axis.title.x = element_text(size=6),
    axis.title.y = element_text(size=6),
    axis.text.x = element_text(size=6),
    axis.text.y = element_text(size=6),
    strip.text.x = element_text(size=6),
    strip.text.y = element_text(angle = 0, size=6),
    legend.position="none",
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank()
    )+
  geom_hline(aes(yintercept = score_adj.y), linetype = 'dashed',
             color = RColorBrewer::brewer.pal(3, "Set2")[1]) 
ggsave(
  file.path(BaseDir,paste0(PlotScore,"PermutedPlots.pdf")),
       width = 15, height = 10, units = "cm"
  )
ggsave(
  file.path(BaseDir,paste0(PlotScore,"PermutedPlots.png")),
       width = 15, height = 10, units = "cm"
  )
  
  
```



Scatter plots - Precision vs Recall vs # Inferred positives?

```{r}

DFToScatterPlot = data.frame(
  M = EvaluationsDF$M,
  D = EvaluationsDF$D,
  CLrn = EvaluationsDF$CLrn,
  K = EvaluationsDF$K,
  Zsd = EvaluationsDF$Zsd,
  score_name = EvaluationsDF$score_name,
  sim_iteration = EvaluationsDF$sim_iteration,
  fctrzn_type = EvaluationsDF$fctrzn_type,
  score_adj = EvaluationsDF$score_adj
)

DFToScatterPlot = DFToScatterPlot[DFToScatterPlot$M == 3
                                  & DFToScatterPlot$D==2000
                                  & is.element(DFToScatterPlot$fctrzn_type, c("MOTL", "Direct"))
                                  ,]

DFToScatterPlotWide = reshape(
  DFToScatterPlot,
  idvar = c("M", "D", "CLrn","K", "Zsd", "fctrzn_type", "sim_iteration"),
  timevar = "score_name",
  v.names = "score_adj",
  direction="wide"
)

colnames(DFToScatterPlotWide) = stringr::str_replace(colnames(DFToScatterPlotWide),stringr::fixed("score_adj."),"")

## create some factors

DFToScatterPlotWide$fctrzn_type = as.factor(DFToScatterPlotWide$fctrzn_type)
DFToScatterPlotWide$fctrzn_type = relevel(DFToScatterPlotWide$fctrzn_type, ref = "Direct")

## make more informative for plotting

DFToScatterPlotWide$D = paste0('D: ',DFToScatterPlotWide$D)
DFToScatterPlotWide$K = paste0('K: ',DFToScatterPlotWide$K)
DFToScatterPlotWide$Zsd = paste0('Zsd: ',DFToScatterPlotWide$Zsd)
DFToScatterPlotWide$CLrn = paste0('CLrn: ',DFToScatterPlotWide$CLrn)

## plot

ggplot(DFToScatterPlotWide, 
       mapping = aes(x=Precision, y=Recall))+ 
  scale_x_continuous(labels = label_comma(accuracy = 0.2)) +
  scale_y_continuous(labels = label_comma(accuracy = 0.2)) +
  geom_point(aes(fill=fctrzn_type, size=Inf_positives), shape = 21, colour = "black") +
  facet_grid(CLrn~K+Zsd, scales = 'free_y') +
  theme_bw()+
  theme(
    strip.text.y = element_text(angle = 0),
    panel.grid.major=element_blank(),
    panel.grid.minor=element_blank()
    )+
  scale_fill_brewer(palette = "Set2")
ggsave(file.path(BaseDir,"AllScoresPlots.png"),
       width = 20, height = 10, units = "cm")

```


Table of median or mean scores

```{r}
TableScore = "F1"
# TableScore = "FitSeconds"

SmryStats = do.call(data.frame, 
                       aggregate(score_adj ~ YDists+D+CLrn+K+Zsd+fctrzn_type, 
                               EvaluationsDF[EvaluationsDF$score_name==TableScore,],
                               # FUN = "quantile", probs = 0.5)
                               FUN = "mean")
                     )

SmryStatsWide = reshape(
  SmryStats,
  idvar = c("YDists", "D", "CLrn","K", "Zsd"),
  timevar = "fctrzn_type",
  v.names = "score_adj",
  direction="wide"
)

colnames(SmryStatsWide) = 
  substr(colnames(SmryStatsWide),
         unlist(gregexpr(pattern = "\\.", text= colnames(SmryStatsWide))) + 1,
         nchar(colnames(SmryStatsWide)))

SmryStatsWide = SmryStatsWide[
  order(SmryStatsWide[,1],SmryStatsWide[,2],SmryStatsWide[,3],SmryStatsWide[,4],SmryStatsWide[,5]),
]

write.table(SmryStatsWide[,-1],
            file.path(BaseDir,paste0(TableScore,"Table.csv")),
            quote = FALSE, sep=',', na='', row.names = FALSE, col.names = TRUE)

```


Fit a gls model with nlme

```{r}

## set up data frame and convert to factors

PlotScore = "F1"

DFToFit = EvaluationsDF[EvaluationsDF$score_name==PlotScore 
                        # & EvaluationsDF$fctrzn_type != "Random_TL"
                        & is.element(EvaluationsDF$fctrzn_type, c("MOTL", "Direct"))
                         & EvaluationsDF$M == 3
                         & EvaluationsDF$D==2000
                         ,]


DFToFit$fctrzn_type = as.factor(DFToFit$fctrzn_type)
DFToFit$fctrzn_type = relevel(DFToFit$fctrzn_type, ref = "Direct")

DFToFit$CLrn = as.factor(DFToFit$CLrn)
DFToFit$K = as.factor(DFToFit$K)
DFToFit$Zsd = as.factor(DFToFit$Zsd)
DFToFit$D = as.factor(DFToFit$D)

DFToFit$sim_id = as.factor(paste0(DFToFit$SimConfig,"_",DFToFit$sim_iteration))
DFToFit$SimConfig = as.factor(DFToFit$SimConfig)

DFToFit$score_to_fit = DFToFit$score_adj


# check the data frame

# summary(DFToFit)

## fit gls with symmetric compound correlation structure based on simulation id

# FctEval_gls = gls(model = score_to_fit ~ fctrzn_type,
#                   data = DFToFit,
#                   correlation = corCompSymm(form = ~1|sim_id)
#                   )

## or with compound correlation structure and heterogeneity of variance 

FctEval_gls = gls(model = score_to_fit ~ fctrzn_type + SimConfig,
                  data = DFToFit,
                  correlation = corCompSymm(form = ~1|sim_id),
                  weights = varIdent(form = ~1|fctrzn_type)
                  )

print(FctEval_gls)

## test contrasts with emmans

FctEval_gls_emm = emmeans(FctEval_gls, "fctrzn_type")
FctEval_gls_ctr = contrast(FctEval_gls_emm, method = "pairwise", adjust = "tukey")

print(FctEval_gls_ctr)


```








